---
title: "Phipps_ms156"
author: "PGD"
date: "28/04/2020"
output: word_document # changed to word doc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the analysis for Bodie's mice

# Questions

> Mother focused

1) Is the MB of mothers different at the time of mating?
  
2) Is the MB of mothers different at the time of birth?
  
3) Is the MB of mothers different at the time the pups are sacrificed?
  
4) Is the MB of mothers different over time (gestation, birth, and pup sacrifice)?

> Pup focused 
  
5) Is the MB of pups different at the time of PVM infection (PND7)?

6) Is the MB of pups different in the early post natal period (PND3)?

7) Is the MB of pups different over time (PND3-7)?
  
> Cross fostering focussed

8) What is the effect of cross fostering on the MB of pups at PND17?

First we got the data in

```{r}

# Data input and subsetting

# Get environmental metadata

env <- read.table('../../Data/env_with_metadata.csv', header=TRUE, sep=',', row.names=1) # updated metadata path

# Get OTU table

otu.tmp <- read.table('../../Data/otu_with_tax_1600.csv', header=TRUE, sep=',', row.names=1) # updated otu table path
otu <- t(otu.tmp[,-76]/1600) # transpose and divide by 1600 to get relative abundances

taxonomy <- otu.tmp[,76] # Make a list of OTUs with the respective taxonomy
taxonomy <- as.data.frame(taxonomy)
row.names(taxonomy) <- row.names(otu.tmp)
# why not generalise to?:
# taxonomy <- as.data.frame(otu.tmp[,'taxonomy'])
# colnames(taxonomy) <- 'taxonomy'

row.names(otu) == row.names(env)

# Subset to questions

env1 <- env[env$Q1 == 1,]
env2 <- env[env$Q2 == 1,]
env3 <- env[env$Q3 == 1,]
env4 <- env[env$Q4 == 1,]
env5 <- env[env$Q5 == 1,]
env6 <- env[env$Q6 == 1,]
env7 <- env[env$Q7 == 1,]
env8 <- env[env$Q8 == 1,]

otu1 <- otu[env$Q1 == 1,]
otu2 <- otu[env$Q2 == 1,]
otu3 <- otu[env$Q3 == 1,]
otu4 <- otu[env$Q4 == 1,]
otu5 <- otu[env$Q5 == 1,]
otu6 <- otu[env$Q6 == 1,]
otu7 <- otu[env$Q7 == 1,]
otu8 <- otu[env$Q8 == 1,]

```

# Load libraries and source custom functions

```{r}
source("../Functions/Functions.R") # updated functions path
library(sciplot)
```


Then we looked at Alpha diversity

```{r}
### Alpha diversity

library(vegan)
library(sciplot)

# Q1
anova(lm(Chao1 ~ factor(Diet), data=env1))
anova(lm(Sobs ~ factor(Diet), data=env1))
#anova(lm(Simp ~ factor(Diet), data=env1))
anova(lm(Shan ~ factor(Diet), data=env1))
anova(lm(PD ~ factor(Diet), data=env1))

# Q2
anova(lm(Chao1 ~ factor(Diet), data=env2))
anova(lm(Sobs ~ factor(Diet), data=env2))
#anova(lm(Simp ~ factor(Diet), data=env2))
anova(lm(Shan ~ factor(Diet), data=env2))
anova(lm(PD ~ factor(Diet), data=env2))

# Q3
anova(lm(Chao1 ~ factor(Diet), data=env3))
anova(lm(Sobs ~ factor(Diet), data=env3))
#anova(lm(Simp ~ factor(Diet), data=env3))
anova(lm(Shan ~ factor(Diet), data=env3))
anova(lm(PD ~ factor(Diet), data=env3))

# Q4
anova(lm(Chao1 ~ factor(Diet) * factor(Time), data=env4))
anova(lm(Sobs ~ factor(Diet) * factor(Time), data=env4))
#anova(lm(Simp ~ factor(Diet) * factor(Time), data=env4))
anova(lm(Shan ~ factor(Diet) * factor(Time), data=env4))
anova(lm(PD ~ factor(Diet) * factor(Time), data=env4))

# Q5
anova(lm(Chao1 ~ factor(Diet), data=env5))
anova(lm(Sobs ~ factor(Diet), data=env5))
#anova(lm(Simp ~ factor(Diet), data=env5))
anova(lm(Shan ~ factor(Diet), data=env5))
anova(lm(PD ~ factor(Diet), data=env5))

# Q6
anova(lm(Chao1 ~ factor(Diet), data=env6))
anova(lm(Sobs ~ factor(Diet), data=env6))
#anova(lm(Simp ~ factor(Diet), data=env6))
anova(lm(Shan ~ factor(Diet), data=env6))
anova(lm(PD ~ factor(Diet), data=env6))

# Q7
anova(lm(Chao1 ~ factor(Diet) * factor(Time), data=env7))
anova(lm(Sobs ~ factor(Diet) * factor(Time), data=env7))
#anova(lm(Simp ~ factor(Diet) * factor(Time), data=env7))
anova(lm(Shan ~ factor(Diet) * factor(Time), data=env7))
anova(lm(PD ~ factor(Diet) * factor(Time), data=env7))

# Q8
anova(lm(Chao1 ~ factor(Cross.foster), data=env8))
anova(lm(Sobs ~ factor(Cross.foster), data=env8))
#anova(lm(Simp ~ factor(Cross.foster), data=env8))
anova(lm(Shan ~ factor(Cross.foster), data=env8))
anova(lm(PD ~ factor(Cross.foster), data=env8))



#Tukey's Honest Significant Difference Post hoc's

# Q4
TukeyHSD(aov(Chao1 ~ factor(Diet) * factor(Time), data=env4))
TukeyHSD(aov(Sobs ~ factor(Diet) * factor(Time), data=env4))
#anova(lm(Simp ~ factor(Diet) * factor(Time), data=env4))
TukeyHSD(aov(Shan ~ factor(Diet) * factor(Time), data=env4))
TukeyHSD(aov(PD ~ factor(Diet) * factor(Time), data=env4))

# Q7
TukeyHSD(aov(Chao1 ~ factor(Diet) * factor(Time), data=env7))
TukeyHSD(aov(Sobs ~ factor(Diet) * factor(Time), data=env7))
#anova(lm(Simp ~ factor(Diet) * factor(Time), data=env7))
TukeyHSD(aov(Shan ~ factor(Diet) * factor(Time), data=env7))
TukeyHSD(aov(PD ~ factor(Diet) * factor(Time), data=env7))

# Q8
TukeyHSD(aov(Chao1 ~ factor(Cross.foster), data=env8))
TukeyHSD(aov(Sobs ~ factor(Cross.foster), data=env8))
#anova(lm(Simp ~ factor(Cross.foster), data=env8))
TukeyHSD(aov(Shan ~ factor(Cross.foster), data=env8))
TukeyHSD(aov(PD ~ factor(Cross.foster), data=env8))


# Plots


# Q1
bargraph.CI(x.factor = factor(Diet), response = Chao1, data=env1, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, data=env1, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, data=env1, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, data=env1, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, data=env1, legend = TRUE)

# Q2
bargraph.CI(x.factor = factor(Diet), response = Chao1, data=env2, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, data=env2, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, data=env2, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, data=env2, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, data=env2, legend = TRUE)

# Q3
bargraph.CI(x.factor = factor(Diet), response = Chao1, data=env3, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, data=env3, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, data=env3, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, data=env3, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, data=env3, legend = TRUE)

# Q4
bargraph.CI(x.factor = factor(Diet), response = Chao1, group = Time, data=env4, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, group = Time, data=env4, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, group = Time, data=env4, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, group = Time, data=env4, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, group = Time, data=env4, legend = TRUE)

# Q5
bargraph.CI(x.factor = factor(Diet), response = Chao1, data=env5, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, data=env5, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, data=env5, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, data=env5, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, data=env5, legend = TRUE)

# Q6
bargraph.CI(x.factor = factor(Diet), response = Chao1, data=env6, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, data=env6, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, data=env6, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, data=env6, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, data=env6, legend = TRUE)

# Q7
bargraph.CI(x.factor = factor(Diet), response = Chao1, group = Time, data=env7, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Sobs, group = Time, data=env7, legend = TRUE)
#bargraph.CI(x.factor = factor(Diet), response = Simp, group = Time, data=env7, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = Shan, group = Time, data=env7, legend = TRUE)
bargraph.CI(x.factor = factor(Diet), response = PD, group = Time, data=env7, legend = TRUE)

# Q8
bargraph.CI(x.factor = factor(Cross.foster), response = Chao1, data=env8, legend = TRUE)
bargraph.CI(x.factor = factor(Cross.foster), response = Sobs, data=env8, legend = TRUE)
#bargraph.CI(x.factor = factor(Cross.foster), response = Simp, data=env8, legend = TRUE)
bargraph.CI(x.factor = factor(Cross.foster), response = Shan, data=env8, legend = TRUE)
bargraph.CI(x.factor = factor(Cross.foster), response = PD, data=env8, legend = TRUE)

```

...and then beta diversity, which first needed some custom functions

```{r}
#  _____                 _   _                 
# |  ___|   _ _ __   ___| |_(_) ___  _ __  ___ 
# | |_ | | | | '_ \ / __| __| |/ _ \| '_ \/ __|
# |  _|| |_| | | | | (__| |_| | (_) | | | \__ \
# |_|   \__,_|_| |_|\___|\__|_|\___/|_| |_|___/
#
# Paul Dennis 

## Axis percent 
# Returns the percentage varation on ordination axes

axis.percent <- function(ordination){
  round((100*eigenvals(ordination)[1:2]/ordination$tot.chi[[1]]),digits=2)
}

# END



## pcoa
# Makes a PCoA ordination object
pcoa <- function(ord){
  cmdscale(dist(ord,'euc'), k=(nrow(ord)-1), eig=TRUE)
}

## axis.percent.pcoa 
# Returns the percentage varation on ordination axes
axis.percent.pcoa <- function(ordination){
  round(as.vector(eigenvals(ordination)/sum(eigenvals(ordination)))[1:2]*100,digits=2)
}

## custom.plot.pcoa
# Plots a PCoA ordination with group enrichment

custom.plot.pcoa <- function(ord,group,title,plottype){
  ordiplot(scores(ord, choices=c(1,2)), type=plottype, 
           xlab=paste("PCoA1 (",axis.percent.pcoa(
             ord)[[1]],"%)",sep=""),
           ylab=paste("PCoA2 (",axis.percent.pcoa(ord)[[2]],"%)",sep=""),
           main=title)
  ordiellipse(ord,group,
              kind='sd')
  #conf=0.95)
  points(ord$points, pch=21, cex=3, bg=factor(group))    
  abline(h=0, lty=3)
  abline(v=0, lty=3)
  legend("topright",legend=unique(factor(group)),pch=19,col=unique(factor(group)))
}


```


```{r}
### Beta Diversity

library(vegan)

# Q1
adonis(sqrt(otu1) ~ factor(Diet), data = env1, method='euc')
pcoa1 <- pcoa(sqrt(otu1))
custom.plot.pcoa(pcoa1,group = factor(env1$Diet), plottype = "t", title = "Q1")

# Q2
adonis(sqrt(otu2) ~ factor(Diet), data = env2, method='euc')
pcoa2 <- pcoa(sqrt(otu2))
custom.plot.pcoa(pcoa2,group = factor(env2$Diet), plottype = "t", title = "Q2")

# Q3
adonis(sqrt(otu3) ~ factor(Diet), data = env3, method='euc')
pcoa3 <- pcoa(sqrt(otu3))
custom.plot.pcoa(pcoa3,group = factor(env3$Diet), plottype = "t", title = "Q3")

# Q4
adonis(sqrt(otu4) ~ factor(Diet) * factor(Time), data = env4, method='euc')
pcoa4 <- pcoa(sqrt(otu4))
env4$Diet.Time = paste(env4$Diet,env4$Time, sep="." )
custom.plot.pcoa(pcoa4,group = factor(env4$Diet.Time), plottype = "t", title = "Q4")

# Q5
adonis(sqrt(otu5) ~ factor(Diet), data = env5, method='euc')
pcoa5 <- pcoa(sqrt(otu5))
custom.plot.pcoa(pcoa5,group = factor(env5$Diet), plottype = "t", title = "Q5")

# Q6
adonis(sqrt(otu6) ~ factor(Diet), data = env6, method='euc')
pcoa6 <- pcoa(sqrt(otu6))
custom.plot.pcoa(pcoa6,group = factor(env6$Diet), plottype = "t", title = "Q6")

# Q7
adonis(sqrt(otu7) ~ factor(Diet) * factor(Time), data = env7, method='euc')
pcoa7 <- pcoa(sqrt(otu7))
env7$Diet.Time = paste(env7$Diet,env7$Time, sep="." )
custom.plot.pcoa(pcoa7,group = factor(env7$Diet.Time), plottype = "t", title = "7")

# Q8
adonis(sqrt(otu8) ~ factor(Cross.foster), data = env8, method='euc')
pcoa8 <- pcoa(sqrt(otu8))
custom.plot.pcoa(pcoa8,group = factor(env8$Cross.foster), plottype = "t", title = "Q8")

```

